Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes

### 2020-08-10 Ben_Manuscripts/hdphmm/response_to_reviewers.tex

    Author reply: We thank the reviewer for their thoughtful assessment of our work.

    %r1_comment2
    \item \textit {Unfortunately, because there is no comparison with alternative techniques, it is 
    very difficult to evaluate this new approach relative to previous work that has examined similar 
    issues. There exist numerous other methods for extrapolating MSDs to longer time, which the 
    authors have not discussed. One such method is the “trajectory-extending kinetic Monte Carlo” 
    (TEKMC) approach of Brown and coworkers see e.g. Macromolecules 2010, 43, 9210-9214, 
    Macromolecules 2015, 48, 7346-7358, and J. Phys Chem. B, 2012, 116, 95-103. The authors should 
    add a comprehensive discussion of how their approach compares to this and similar methods. The 
    paper would be significantly improved if the authors were to perform and present TEKMC results 
    for their system, so that one could better evaluate the performance of their proposed method.}

    Author reply: We thank the reviewer for bringing our attention to these other interesting 
    approaches. They certainly merit discussion as alternative approaches for estimating mean 
    squared displacements on long timescales. We have added some discussion of these methods to 
    the introduction.

    \begin{quote}
    A number of approaches have been explored to inexpensively extend the relatively short
    trajectories output by molecular simulations which exhibit complex diffusive behaviors
    onto much longer timescales. The trajectory extending kinetic Monte Carlo (TEKMC) approach
    divides a unit cell into a three dimensional grid and quantifies the transition 
    probabilities of particles that can move between each subdivision allowing one to generate
    stochastic trajectories that reach the diffusive regime~\cite{neyertz_trajectory-extending_2010}
    If the dynamics of the system are fast enough, it is possible to parameterize TEKMC models with
    relatively short simulation trajectories which facilitates systematic studies of small perturbations
    to material constituents.~\cite{hanson_computer_2012}
    In a related approach, dynamic bond percolation (DBP) theory can be used to quantify the
    rate of transition between sites in a dynamically disordered medium and long timescale
    behavior can be propagated using kinetic Monte Carlo simulations.~\cite{druger_dynamic_1983}
    Webb et al. extended this approach by defining sites based on the molecular details of 
    a polymer electrolyte system, thus allowing them to explain ion diffusivities predicted by
    the model in terms of chemically specific interactions.~\cite{webb_chemically_2015}
    \end{quote}

@@

    We have also modified the first sentence of the paragraph that follows in order to maintain
    good flow.

    Original text:
    \begin{quote}
    We have already advanced our understanding of LLC monomer design in our previous work
    by building stochastic time series models based on the chemical intuition gained
    from qualitative MD studies of solute transport in an H\textsubscript{II} phase LLC
    membrane.
    \end{quote}

    Modified text:
    \begin{quote}
    Based on the chemical intuition gained from our previous qualitative MD studies of 
    solute transport in an H\textsubscript{II} phase LLC membrane, we have contributed our
    own stochastic modeling approaches with goals similar to both TEKMC and DBP.
    \end{quote}

    %BJC2: 
    We agree that a full comparison of our method to other methods for computing long timescale 
    diffusion such as TEKMC would be very interesting and instructive but feel it would be better 
    suited as a separate study where we could evaluate a number of approaches, including perhaps
    recent approaches we have developed within our research group.~\cite{coscia_capturing_2020}
    %BJC2: added to make the point about our system being more complicated than those studied.
    It would be illuminating to test these methods on systems such as ours where there is
    a rich diversity of intermolecular interactions that work to immobilize solutes.
    Such a large-scale comparison would take more space and time than would be appropriate for 
    the current paper. Although methods such as TEKMC seems a fairly straightforward approach to
    implement, it is not immediately clear how we could optimally adapt the methodology to a 
    monoclinic unit cell with two dimensional order without in-depth testing.
    % the grids that divide the homogeneous cubic unit cells of the referenced TEKMC studies to 
    % our system which uses a monoclinic unit cell and has 2 dimensional order. 
    %BJC2: maybe push back on the CS-DBP here? 
    Regarding the CS-DBP model, we are not convinced it would be suitable for our goals because
    it relies on defining bound states and modeling hops between them. There are numerous potential
    locations within our LLC system where solutes can become bound in some way. The HDP-AR-HMM
    circumvents the need for relatively complicated methods used to identify binding sites 
    by automatically identifying distinct dynamics characteristic of a class of bound states.

@@

%MRS: additional thoughts:

%1. Dynamic Bond Percolation theory (both the original version and the
%chemically specific one) an and TEKMC are based on spatial clustering
%and coarse-graining. Our approach is based on clustering and
%coarse-graining the dynamics.  Can be more powerful, doesn't assume
%(or minimally assumes - in some cases we assume pore vs. nono-pore)
%which configurations are important, which matters when there are a
%diversity of different types of bound states, and diversity of
%dynamics (in the pore, associated with the head groups, etc).

%2. Dynamical behavior is more complex with our solutes than ion
%interactions because of the richer range of interactions involved
%(ionic binding PLUS hydrophobic binding, i.e. more types of bound
%states; it's not just binding and hopping, and thus something

%BJC2: I believe I had orginally addressed this briefly as well. Feel free to expand if you feel there is more to be said
%3. Configurational behavior is more complex with the nanostructures,
%because one cannot simply divide the structure into cubes
%isotropically.  It may be easier to use our approach with more
%diverse configurations, because we don't need to worry about
%partitioning correctly.  This does push some of the difficulty into
%choosing the coordinate system for the dynamics, but we have a lot
%more freedom about choosing coordinate systems for the model than one
%would using a uniform (or even nonuniform) partitioning of the
%system. Yes, there are difficult choices to make but the model has
%the freedom to make these choices.

@@

    We do agree that there are certain basic simulation details that we have neglected to 
    mention. Therefore we have made the following edits to section 2.1 of the methods:

    We added the highlighted text to the first sentence of the section:
    \begin{quote}
      We studied transport of solutes in the H\textsubscript{II} phase using \hl{our
      previously developed} atomistic molecular model of four pores in a 
      \hl{$9.4\times9.4\times7.6$ nm} monoclinic unit cell with 10\% water by weight, 
      \hl{totaling 62,000 atoms}.    
    \end{quote}

    We have also added the following text describing the model:
    \begin{quote}
    We modeled water molecules using the TIP3P water model,~\cite{jorgensen_comparison_1983}
    We used the Antechamber module shipped with AmberTools18 to parameterize monomers and 
    solutes using the Generalized Amber Force Field (GAFF).~\cite{wang_development_2004} We 
    assigned charges using the AM1-BCC model as implemented in Antechamber.~\cite{jakalian_fast_2000,jakalian_fast_2002}
    GAFF is very commonly used in simulations with diverse organic chemical functionality, 
    and was developed for consistency within the AMBER ecosystem, which uses TIP3P as the
    water model. We chose these standard force fields and generalizable parameterization 
    approaches so that a range of solutes and monomers could be self-consistently and easily 
    studied. The main goal of this work is to develop stochastic models which can reproduce
    solute behavior on MD time scales, independent of the choice of force field. In future 
    work, it may be beneficial to employ specialized force fields, including polarizable 
    force fields or other higher-order effects in order to improve the accuracy of structure 
    parameters and selectivity predictions.
    \end{quote}

    Finally, we have added a figure with a picture of the system which we are studying 
    (see~\ref{fig:membrane_side_profile}).
    \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{membrane_side_profile.pdf}
    \caption{When combined with 10 wt\% water, the LLC monomer studied in this work forms the 
    H\textsubscript{II} phase which is characterized by hexagonally packed columns with 
    hydrophilic pores. In the figure, white and red water molecules mark the membrane pores.
    The monomer fills the rest of the space. Note that water molecules are present in the 
    tails, but they are hidden for the sake of clarity. }\label{fig:membrane_side_profile}
    \end{figure}

@@

    Author reply: We thank the reviewer for making this point. 
    %BJC2: modded based on MRS suggestion    
    In systems with less obstructions to motion, such as LLCs with a higher water content, most 
    solutes and water will be relatively unimpeded, and diffusion constants can be obtained by 
    standard techniques.    
    However, we do contend that our method would still be very useful for studying
    solute dynamics in these higher water content systems 
    %BJC2: incorporated MRS suggestion with small modifications    
    by allowing one to study the dynamics of solutes that have longer residence times and interact
    with moieties in a complex medium. %portions of the system away from the pore center.
    %Interactions characterized by different dynamical behavior even on short time scales can
    %still be highly informative. 
    This is the basis for the point we make in the introduction:``...direct calculations of diffusion 
    constants may obscure the molecular mechanisms controlling transport". Therefore we would
    prefer not to limit the range of application of this method. 
    %BJC2: added    
    However, we do agree that it would be helpful to the reader to clarify that the level of modeling 
    which we present in this work is certainly not merited in cases where they are only interested in
    diffusion constants. Therefore, we have modified a paragraph in the introduction in order to
    encourage readers to pursue simpler approaches if the system allows it.

    Updated text with additions highlighted in yellow:

    \begin{quote}
    In this work, we are particularly interested in the design of lyotropic liquid 
    crystal (LLC) monomers, a class of amphiphilic molecules whose ordered self-assembled
    phases can be cross-linked into mechanically strong membranes capable of highly 
    selective separations. Inverted hexagonal phase, or H\textsubscript{II}, LLC membranes
    are characterized by hexagonally packed, uniform-sized and straight pores, an ideal 
    geometry for high throughput transport. Their pores are lined with the LLC monomer's
    functional groups which can potentially be designed to interact with solutes in a 
    chemically-specific manner. It would be extremely useful to experimentalists if we 
    could connect LLC monomer design with measurable quantities such as diffusivity and
    selectivity. \hl{In many systems, including LLC membranes with water contents higher
    than the 10 wt\% water system studied here, making this connection can be relatively 
    straightforward calculation with standard techniques since small solutes and water 
    can move relatively unimpeded and therefore enter a diffusive regime on timescales 
    accessible to MD. When possible, we encourage the reader to take advantage of the 
    simplest approach possible. However, for the system that has captured our interest,} 
    this goal is complicated by the non-Brownian hopping and trapping behavior induced 
    by the membrane's diverse topology and inhomogeneous structure resulting in subdiffusive
    behavior out to multiple microseconds of simulation.~\cite{coscia_understanding_2019,coscia_chemically_2019}
    \end{quote}

@@

    %r1_comment5
    \item \textit{In the HDP-AR-HMM method, it is not clear to me how the solute molecules are 
    constrained to be confined within the pore radius (which they must be on physical grounds 
    based on steric interactions). From Figure 2, it “looks” like this happens to work out, but
    for more complicated systems (e.g. gyroid LLCs with 3D channels) such a constraint may be 
    necessary and it is unclear how the method would generalize. This should be discussed.}

    Author reply: 
    %We thank the reviewer for posing this question as this a constraint to which 
    %we gave thought but neglected to emphasize in the main text. 
    %BJC: he's not saying that solutes should stay in the aqueous pore region right? We talk pretty
    % extensively about solute behavior in the tails so his steric argument is confusing me.

@@

    We have added the following text to the end of Section 2.3 in order to make it clear that
    the trajectory generation procedure will require more thought for different geometries:
    \begin{quote}
    The parameterization of $\mathbf{c}$, $A$ and $\Sigma$ as presented above has been designed 
    specifically for the cylindrical pore geometry at hand. Our group is actively investigating
    appropriate ways to parameterize more complicated geometries such as bicontinuous cubic phase
    LLC membranes. 
    \end{quote}

    %BJC: figured I'd get your feedback on above before writing the part that will go into the text.

    %r1_comment6
    \item \textit{As stated in the introduction, one motivation for the work is that “...one 
    cannot reliably compute diffusion constants because MSDs are often non-linear on the timescales
    accessible to molecular simulation”. Later on in section 2.4, the authors state “For the 
    purposes of our analysis, we assume ergodicity of the 24 solute trajectories.” These are 
    contradictory statements. If the trajectories are ergodic, then MSDs will be linear by 
    definition The authors should clarify. Similarly, in Figure 7, if the simulations are indeed 
    ergodic the red and blue bars should be proportional to each other.}

    Author reply: We thank the reviewer for identifying the potential confusion that could be
    caused by our use of the word `ergodic'. Figure 7 clearly demonstrates that individual trajectories
    are not by themselves ergodic. Our intention was to communicate our assumption that collectively all 24 
    trajectories visit all possible dynamic states with representative probability. This is 
    necessary in order to make accurate long timescale predictions. The 5 $\mu s$ length of our 
    simulations was intended to increase the validity of this assumption.

### 2020-08-10 Ben_Manuscripts/hdphmm/Draft.tex
Modifications and additions. Addressing requested non-scientific changes

  In this work, we are particularly interested in the design of lyotropic liquid 
  crystal (LLC) monomers, a class of amphiphilic molecules whose ordered self-assembled
  phases can be cross-linked into mechanically strong membranes capable of highly 
  selective separations. Inverted hexagonal phase, or H\textsubscript{II}, LLC membranes
  are synthesized with 10 \% water by weight and are characterized by hexagonally packed,
  uniform-sized and straight pores, an ideal geometry for high throughput transport. 
  Their pores are lined with the LLC monomer's functional groups which can potentially 
  be designed to interact with solutes in a chemically-specific manner. It would be 
  extremely useful to experimentalists if we could connect LLC monomer design with 
  measurable quantities such as diffusivity and selectivity.
  %r1_comment4
  In many systems, including LLC membranes with higher water contents, making this 
  connection is a relatively straightforward calculation since small solutes and water
  can move relatively unimpeded and therefore enter a diffusive regime on timescales
  accessible to MD. When possible, we encourage the reader to take advantage of the
  simplest approach possible.
  However, in the systems of interested to us,
  this goal is complicated by the non-Brownian hopping and trapping behavior induced by the membrane's diverse
  topology and inhomogeneous structure resulting in subdiffusive behavior out to 
  multiple microseconds of simulation.~\cite{coscia_understanding_2019,coscia_chemically_2019}

@@

@@

  \bibliographystyle{ieeetr}
  \bibliography{hdphmm}

>>>>

  In this work, we are particularly interested in the design of lyotropic liquid 
  crystal (LLC) monomers, a class of amphiphilic molecules whose ordered self-assembled
  phases can be cross-linked into mechanically strong membranes capable of highly 
  selective separations. Inverted hexagonal phase, or H\textsubscript{II}, LLC membranes
  are synthesized with 10 \% water by weight and are characterized by hexagonally packed,
  uniform-sized and straight pores, an ideal geometry for high throughput transport (see
  Figure~\ref{fig:membrane_side_profile}). 
  Their pores are lined with the LLC monomer's functional groups which can potentially 
  be designed to interact with solutes in a chemically-specific manner. It would be 
  extremely useful to experimentalists if we could connect LLC monomer design with 
  measurable quantities such as diffusivity and selectivity.
  %r1_comment4
  In many systems, including LLC membranes with water contents higher than the 10 wt\% 
  water system studied here, making this connection can be relatively straightforward 
  calculation with standard techniques since small solutes and water can move relatively
  unimpeded and therefore enter a diffusive regime on timescales accessible to MD. When
  possible, we encourage the reader to take advantage of the simplest approach possible. 
  However, for the system that has captured our interest, this goal is complicated by 
  the non-Brownian hopping and trapping behavior induced by the membrane's diverse
  topology and inhomogeneous structure resulting in subdiffusive behavior out to 
  multiple microseconds of simulation.~\cite{coscia_understanding_2019,coscia_chemically_2019}

  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{membrane_side_profile.pdf}
  \caption{When combined with 10 wt\% water, the LLC monomer studied in this work forms the 
    H\textsubscript{II} phase which is characterized by hexagonally packed columns with 
    hydrophilic pores. In the figure, white and red water molecules mark the membrane pores.
    The monomer fills the rest of the space. Note that water molecules are present in the 
    tails, but they are hidden for the sake of clarity.}\label{fig:membrane_side_profile}
  \end{figure}

@@

  In consideration of our ultimate goal, which is to use high-throughput simulations to 
  facilitate our understanding of LLC membrane design for solute-specific separations, the size 
  of our system and length of the accompanying molecular simulations appears to be somewhat 
  impractical. Dependent on one's goals, it may not be necessary to run such long simulations in order 
  to extract valuable mechanistic insight. We originally uncovered the mechanisms discussed
  in this study using trajectories a fraction of the length here.\cite{coscia_chemically_2019} A 
  strong mechanistic understanding of solute transport with respect to the chemical features of a 
  given LLC monomer may be enough for experimentalists to eliminate entire subsets of potential
  design changes. However, it may still be valuable to consider how one might make these simulations
  amenable to high-throughput studies. We would like to emphasize that our large, fully atomistic 
  models will be extremely valuable as a benchmark system used to explore less expensive models. 
  For example, we could explore systems built with fewer LLC monomers by making the membrane pores
  shorter. We could also build systems with higher concentrations of solutes in order to gather more 
  dynamical data on shorter timescales. Using what we have learned about the radial distribution
  functions of these solutes in this and our previous work~\cite{coscia_chemically_2019}, it may be possible to 
  learn how to create initial configurations with solutes initially distributed in a way that
  is close to their equilibrium distribution, allowing solutes to thoroughly explore the 
  accessible membrane structural space more quickly and reducing the amount of unequilibrated
  data that must be discarded. Alternatively, we could use grand canonical Monte Carlo simulations
  to load solutes into the membrane~\cite{snurr_prediction_1993}. Finally, it may be possible 
  to speed up our simulations by reducing the number of atoms using united-atom representations
  of the liquid crystal monomers~\cite{horta_gromos-compatible_2016} or by using hydrogen mass 
  repartitioning to allow us to increase the MD time step.~\cite{hopkins_long-time-step_2015}

@@

  % \bibliographystyle{ieeetr}
  \bibliography{hdphmm}
